---
title: "SIMD 2020 plot"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Plot SIMD 2020 data by local authority. Aiming to visualise the spread of deprivation within each council. 

##Packages

Install necesarry packages and set base plotting theme.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(readxl)
library(janitor)
theme_set(theme_minimal(base_size = 26, base_family = "Roboto") +
            theme(panel.grid.minor = element_blank()))
```


## Data

Data comed from the Scottish Government's SIMD 2020 webpage [here](insert link address) 

```{r}
#The third sheet in the csv is the one with the data that I want. 
simd <- read_excel("raw_data/SIMD_2020_Datazone_lookup_tool.xlsx", sheet = 3)
```


# By Local Authority

## Tidy up

A few things required to get this data into a format ready for plotting. Comments in the below chunk describe what is happening at each stage. 

```{r}
simd %>% 
  #convert all column names to lower and snake case
  #I should come up with better names here but am being lazy
  janitor::clean_names() %>% 
  #keep only the columns of interest
  select(l_aname, population, simd2020_decile) %>% 
  #clean_names didn't quite have the desired effct for la name - sort it here
  rename(la_name = l_aname) %>% 
  #Now group by council and simd decile
  group_by(la_name) %>% 
  #Add a variable which counts the number of datazones
  #(in each council)
  mutate(n_dz = n()) %>% 
  #now group by la and SIMD decile
  group_by(la_name, simd2020_decile) %>%
  #Now calculate pct of datazones in each decile per LA
  #P.S. don't multiply by 100 here as ggplot will do that for us later
  mutate(pct = n()/n_dz) %>% 
  #easier to read (for a human it breaking the pipe here for a look)
  arrange(la_name, simd2020_decile) %>% 
  ungroup %>% 
  #keep the first row of data for each council and decile 
  #(basically a dedup)
  distinct(la_name, simd2020_decile, .keep_all = TRUE) %>% 
  #now factorise the SIMD variable for better plotting options
  mutate(simd2020_decile = factor(simd2020_decile,
                                  levels = c(1:10),
                                  labels = c("1", "2", "3", "4", "5",
                                             "6", "7", "8", "9", "10"))) %>% 
  #drop the now unnecessary population variable
  select(-population) -> clean_simd
#have a look at the outcome
clean_simd
```


## Plot

All ready to plot now...

```{r, fig.width=20, fig.height=18}
clean_simd %>% 
  ggplot(aes(simd2020_decile, pct, fill = simd2020_decile)) +
  geom_col() +
  scale_fill_ptol(guide = guide_legend(nrow = 1),
                  labels = c("1-most deprived", "2", "3", "4", "5",
                             "6", "7", "8", "9", "10-most affluent")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~la_name, strip.position = "bottom") +
  theme(legend.position = "bottom",
        axis.text.x = element_blank()) +
  labs(title = "Percentage of datazones in SIMD 2020 deciles",
       subtitle = "by Local Authority",
       x = "",
       y = "",
       fill = "SIMD decile", 
       caption = "Data from https://www.gov.scot/publications/scottish-index-of-multiple-deprivation-2020-data-zone-look-up/") -> simd_la
simd_la
```

## Save

```{r}
ggsave("plots/simd_2020_plot.png", simd_la, width = 20, height = 18)
```

## Tale of two cities plot

```{r fig.width=16, fig.height=9}
clean_simd %>% 
  filter(la_name %in% c("Glasgow City", "City of Edinburgh")) %>% 
  ggplot(aes(simd2020_decile, pct, fill = simd2020_decile)) +
  geom_col() +
  scale_fill_ptol(guide = guide_legend(nrow = 1),
                  labels = c("1-most deprived", "2", "3", "4", "5",
                             "6", "7", "8", "9", "10-most affluent")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~la_name, strip.position = "bottom") +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.caption = element_text(size = 12)) +
  labs(title = "Percentage of datazones in SIMD 2020 deciles",
       subtitle = "A tale of two cities",
       x = "",
       y = "",
       fill = "SIMD decile", 
       caption = "Data from https://www.gov.scot/publications/scottish-index-of-multiple-deprivation-2020-data-zone-look-up/") -> simd_la_2cities
simd_la_2cities
```

```{r}
ggsave("plots/simd_2020_2cities.png", simd_la_2cities, width = 16, height = 9)
```

## Save

# By Health Board

## Tidy up

```{r}
simd %>% 
  janitor::clean_names() %>% 
  select(h_bname, population, simd2020_decile) %>% 
  rename(hb_name = h_bname) %>% 
  group_by(hb_name) %>% 
  mutate(n_dz = n()) %>% 
  group_by(hb_name, simd2020_decile) %>% 
  mutate(pct = n()/n_dz) %>% 
  arrange(hb_name, simd2020_decile) %>% 
  ungroup %>% 
  distinct(hb_name, simd2020_decile, .keep_all = TRUE) %>% 
  mutate(simd2020_decile = factor(simd2020_decile,
                                  levels = c(1:10),
                                  labels = c("1", "2", "3", "4", "5",
                                             "6", "7", "8", "9", "10"))) %>% 
  select(-population) -> hb_simd
```

## Plot


```{r, fig.width=20, fig.height=18}
hb_simd %>% 
  ggplot(aes(simd2020_decile, pct, fill = simd2020_decile)) +
  geom_col() +
  scale_fill_ptol(guide = guide_legend(nrow = 1),
                  labels = c("1-most deprived", "2", "3", "4", "5",
                             "6", "7", "8", "9", "10-most affluent")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  facet_wrap(~hb_name, strip.position = "bottom") +
  theme(legend.position = "bottom",
        axis.text.x = element_blank()) +
  labs(title = "Percentage of datazones in SIMD 2020 deciles",
       subtitle = "by Health Board",
       x = "",
       y = "",
       fill = "SIMD decile", 
       caption = "Data from https://www.gov.scot/publications/scottish-index-of-multiple-deprivation-2020-data-zone-look-up/") -> simd_hb
simd_hb
```

## Save

```{r}
ggsave("plots/simd_2020_plot_hb.png", simd_hb, width = 16, height = 9)
```

*Polar plots*???




# Session Info

```{r}
devtools::session_info()
```

